Classless Java: Tuning Java Interfaces

### 2015-12-02 paper/sections/Formalization.tex
Revising formalization.

Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
\noindent$\begin{array}{l}
\InTextDef{15ex}{
\mBody(\m,\C_0)
}{
\override(\methods(\m),
\shadow(\tops(\m,\Cs\ )))
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\cC}
\end{array}$

\noindent The definition of $mBody$ delegates the work to three others auxiliary functions:
 $\tops(m,\Cs)$, $\shadow(\methods)$ and $\override(\method,\method')$.

\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
the non static ones that are not transitively reachable by following another, less specific, superinterface chain. Formally:

\noindent$\begin{array}{l}
\InTextDef{22ex}{
\method\in\tops(\m,\Cs)}{
}{
\mBody(\m,\C)=\method
}\\
\InTextWith{
 \method\mbox{ not a static method, and }
 \C\in\Cs\ \mbox{ such that }
\forall \C'\in\Cs\setminus\C \mbox{ not } \C' \subtype \C}

\end{array}$

@@

\paragraph{override} models how the implementation in an interface can
override implementation in the superinterface, even in case of a
conflict.  Note how we use the special value $\none$, and how (forth
case) overriding can solve a conflict.

>>>>

Defining \mBody{} is not trivial, and requires quite a lot of attention to the specific model of Java Interfaces, and on how it differs w.r.t. Java Class model.
$\mBody(\m,\C)$ denotes the actual method $\m$ (body included) that
interface $\C$ owns. The method can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.
\noindent$\begin{array}{l}
\InTextDef{15ex}{
\mBody(\m,\C_0)
}{
\override(\methods(\m),
\shadow(\tops(\m,\C_0)))
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}
\end{array}$

\noindent The definition of $\mBody$ delegates the work to three others auxiliary functions:
 $\tops(m,\C_0)$, $\shadow(\methods)$ and $\override(\method,\method')$.

\paragraph{\tops{}} recovers from the interface table only the ``needed'' methods, that is,
the non-static ones that are not transitively reachable by following another, less specific, superinterface chain.
\hl{Another two functions $\super(\C_0)$ and $\get(\m,\C_0)$ are used for the definition. $\super(\C_0)$ collects the
super interfaces of $\C_0$ recursively, and $\get(\m,\C_0)$ returns the non-static method declared in $\C_0$ with name $\m$, or $\emptyset$ otherwise.
Finally, $\tops(\m,\C_0)$ returns the topmost methods $m$ from the super interfaces.}
Formally:

\noindent$\begin{array}{l}
\InTextDef{13ex}{
\tops(\m,\C_0)}{
}{
\{\ \get(\m, \C)\ |\ \C\in\super(\C_0), \get(\m, \C) \ne\emptyset,\mbox{ and }
}\\
\hspace{2.1in}\forall\C'\in\super(\C_0)\mbox{ and }\get(\m,\C')\ne\emptyset: \C'<:\C\Rightarrow\C'=\C\ \}\\
\InTextDef{13ex}{
\super(\C_0)}{
}{
\{\ \C'\ |\ \C'\in\Cs,\mbox{ or }\exists\C\in\Cs,\C'\in\super(\C)\ \}
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC}\\
\InTextDef{13ex}{
\get(\m,\C_0)}{
}{
\method
}\\
\InTextWith{\metaVar{IT}(\C_0) =
\ \ann\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC\methods\ \cC, \method\in\methods\mbox{ not static},\m=\m^\method}
\end{array}$

@@

\paragraph{override} models how the implementation in an interface can
override implementation in the superinterface, even in case of a
conflict.  Note how we use the special value $\none$, and how (forth
case) overriding can solve a conflict. \hl{The notation in the last case below
means $\method'$ is not static, and is supertype of $\method$.}

### 2015-12-02 paper/sections/Translation.tex
Revising formalization.

\haoyuan{I think $\QM{with#}\m\notin\dom(\methods)$ is not enough. The withM method with return type $I_0$ can also be defined already in a super interface. Like:
\[\begin{array}{l}\QM{
interface A}\ \oC\QM{int x(); B withX(int x);}\cC\\
\mixinAnn\ \QM{interface B extends A}\ \oC\cC\end{array}\]}
\marco{It is ok, in that case we just regenerate it identical}

\noindent$\begin{array}{ll}
\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{
 \isWith(\mBody(\QM{with#}\m, \C_0), \C_0),}\\
\tab\tab{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
\tab\tab{
\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM_\m\notin\dom(\methods)}\\
\end{array}$

The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods \hl{directly declared in} $\C_0$.

>>>>

\noindent$\begin{array}{ll}
\InTextDef{16em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{
 \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
\hspace{2.65in}{
\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
\InTextDef{16em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0),}\\
\hspace{2.65in}{
\mbox{ and } \QM_\m\notin\dom(\methods)}\\
\end{array}$

The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$, \hl{since overloading is not modelled}.
